Stateless Model Checking with Data-Race Preemption Points

### 2016-03-23 oopsla-16/evaluation.tex
Back up shaky partial verif results with a citation to sealing.
minor
Fill in icb column on table. Back out variance paragraph. some other commence

\revision{
%{\bf Variance.}
% TODO CAMREADY: Probably will need at least a 3rd run of dis.
Because \quicksand~can be nondeterministic in its scheduling of state spaces,
we ran the QS-DR tests a second time to quantify the variance.
We found the area between the two QS-DR curves to be
% TODO
{\bf XX\%}
of the area between QS-DR and SSS-MC-ICB (the next closest),
% TODO cite
which we consider an insignificant amount of variance.
}

@@

% TODO: fix this table.
\begin{table*}[t]
	\begin{center}
		\small
	\begin{tabular}{r|c||c|c|c|c|c|c||c|c||c|c}
		% TODO: Update SSSMC column to have more favourable ICB numbers.
		& {\bf Num.} & {\bf DPOR} & {\bf ICB} & {\bf QS} & {\bf DR} & {\bf Nondet.} & {\bf Malloc} & {\bf Mutual} & {\bf Avg. tested} & {\bf Total} & {\bf Untested} \\
		{\bf Test} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR bugs} & {\bf DRs} & {\bf timeouts} & {\bf subset SSes} & {\bf DR PPs} & {\bf DR PPs} \\
		\hline
		% FIXME CAMREADY: update FRMs
					% TODO: update to be ShMem not DPOR
											% TODO: recount with ICB
				% #test  sssmc   icb      qs      dronly  nondets  FRMs   timeout comp.SSes DRPPs untested DRPPs
		{\tt bcast}	& 79	& 7	& TODO	& 11	& 4	& 4	& 51	& 0	& -	& 912	& 107	\\
		{\tt join} 	& 79	& 14 	&	& 24	& 10	& 6	& 333	& 5	& 60.8	& 781	& 292	\\
		{\tt mx} 	& 79	& 1	&	& 10	& 9	& 1	& 7	& 0	& -	& 829	& 1	\\
		{\tt sem} 	& 79	& 10	&	& 16	& 6	& 5	& 140	& 29	& 73.3	& 753	& 279	\\
		{\tt signal} 	& 79	& 4	&	& 10	& 7	& 6	& 180	& 10	& 54.2	& 1118	& 391	\\
		{\tt rwlock} 	& 79	& 26	&	& 28	& 2	& 2	& 125	& 19	& 29.6	& 915	& 634	\\
		\hline
		{\tt sched} 	& 59	& 1	&	& 7	& 6	& 4	& 0	& 0	& -	& 144	& 3	\\
		{\tt alarm} 	& 44	& 5	&	& 21	& 17	& 3	& 35 	& 22	& 9.1	& 115	& 89	\\
		{\tt wait} 	& 52	& 23	&	& 30	& 7	& 1	& 31	& 8	& 28.0	& 142	& 31	\\
		\hline
		{\bf Total}	& 629	& 91	&	& 157	& 68	& 32	& 902	& 93	& 42.6 	& 5709	& 1827	\\
	\end{tabular}
	\end{center}
	\caption{Summary of bugs and data races found by each test program.
		% TODO
		DPOR is the control; QS is \quicksand.
		``DR bugs'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}});
	among those, ``Nondeterministic DR bugs'' counts how many candidates required MC integration to identify (\sect{\ref{sec:eval-dr}}).
	``Malloc DRs'' counts how many malloc-recycle false positives we suppressed (\sect{\ref{sec:eval-dr}}).
	``Mutual timeouts'' counts how often both SSS-MC and \quicksand~timed out with no bug found;
	among those, ``Average tested subset SSes'' counts how many partial verifications \quicksand~provided on average for each test (\sect{\ref{sec:eval-sssmc}}).
	``Total DR PPs'' counts how many unique data-racing instructions we identified among tests where we found no bugs;
	among those, ``Untested DR PPs'' counts how many could not be checked in the time limit (\sect{\ref{sec:future}}).
		}
	\label{tab:drbugs}
\end{table*}

@@

{\bf Partial verification.}
When a MC job times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
%Beyond using the state space estimator to guess the percent coverage,
We know of no approach to quantify the probability
that a bug would be exposed by an untested interleaving,
but \quicksand~at least reports which subsets of PPs resulted in state spaces that did complete in time.
%\quicksand~does not attempt any such quantitative guarantee,
%but can at least report
On 129 tests, the SSS-MC control experiment timed out after 10 hours with no bugs found.
Among these tests, \quicksand~found bugs in 36.
For the other 93, we show the number of state spaces \quicksand~was able to complete in the ``Average tested subset SSes'' column of Table~\ref{tab:drbugs}.
%and the number of unique PPs among them.
%In 6 cases, \quicksand~also failed to complete anything; beyond these,
%between 1 and 253 state spaces were completed for each test
%Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.

@@

>>>>

% TODO CAMREADY
%\revision{
%%{\bf Variance.}
%% TODO CAMREADY: Probably will need at least a 3rd run of dis.
%Because \quicksand~can be nondeterministic in its scheduling of state spaces,
%we ran the QS-DR tests a second time to quantify the variance.
%% TODO CAMREADY
%% We found the maximum area between any two QS-DR curves to be
%We found the area between the two QS-DR curves to be
%% TODO
%{\bf XX\%}
%of the area between QS-DR and SSS-MC-ICB (the next closest),
%% TODO cite
%which we consider an insignificant amount of variance.
%}

@@

% TODO: fix this table.
\begin{table*}[t]
	\begin{center}
		\small
	\begin{tabular}{r|c||c|c|c|c|c|c||c|c||c|c}
		% TODO: Update SSSMC column to have more favourable ICB numbers.
		& {\bf Num.} & {\bf DPOR} & {\bf ICB} & {\bf QS} & {\bf DR} & {\bf Nondet.} & {\bf Malloc} & {\bf Mutual} & {\bf Avg. tested} & {\bf Total} & {\bf Untested} \\
		{\bf Test} & {\bf tested} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf bugs} & {\bf DR bugs} & {\bf DRs} & {\bf timeouts} & {\bf subset SSes} & {\bf DR PPs} & {\bf DR PPs} \\
		\hline
		% FIXME CAMREADY: update FRMs
					% TODO: update to be ShMem not DPOR
											% TODO: recount with ICB
				% #test  sssmc   icb      qs      dronly  nondets  FRMs   timeout comp.SSes DRPPs untested DRPPs
			% TODO: If possible, add a column for how many false negatives (incorrect verif claims) ShMem had.
		{\tt bcast}	& 79	& 7	& 7	& 11	& 4	& 4	& 51	& 0	& -	& 912	& 107	\\
		{\tt join} 	& 79	& 14 	& 14	& 24	& 10	& 6	& 333	& 5	& 60.8	& 781	& 292	\\
		{\tt mx} 	& 79	& 1	& 1	& 10	& 9	& 1	& 7	& 0	& -	& 829	& 1	\\
		{\tt sem} 	& 79	& 10	& 11	& 16	& 6	& 5	& 140	& 29	& 73.3	& 753	& 279	\\
		{\tt signal} 	& 79	& 4	& 4	& 10	& 7	& 6	& 180	& 10	& 54.2	& 1118	& 391	\\
		{\tt rwlock} 	& 79	& 26	& 27	& 28	& 2	& 2	& 125	& 19	& 29.6	& 915	& 634	\\
		\hline
		{\tt sched} 	& 59	& 1	& 1	& 7	& 6	& 4	& 0	& 0	& -	& 144	& 3	\\
		{\tt alarm} 	& 44	& 5	& 16	& 21	& 17	& 3	& 35 	& 22	& 9.1	& 115	& 89	\\
		{\tt wait} 	& 52	& 23	& 23	& 30	& 7	& 1	& 31	& 8	& 28.0	& 142	& 31	\\
		\hline
		{\bf Total}	& 629	& 91	& 104	& 157	& 68	& 32	& 902	& 93	& 42.6 	& 5709	& 1827	\\
	\end{tabular}
	\end{center}
	\caption{Summary of bugs and data races found by each test program.
		% TODO: Add Shared-Mem
		DPOR/ICB are the controls; QS is \quicksand.
		``DR bugs'' counts among \quicksand's bugs how many required data-race PPs to expose (\sect{\ref{sec:eval-sssmc}});
	among those, ``Nondeterministic DR bugs'' counts how many candidates required MC integration to identify (\sect{\ref{sec:eval-dr}}).
	``Malloc DRs'' counts how many malloc-recycle false positives we suppressed (\sect{\ref{sec:eval-dr}}).
	% TODO CAMREADY: count timeouts for ICB instead?
	``Mutual timeouts'' counts how often both SSS-MC-DPOR and \quicksand~timed out with no bug found;
	% TODO CAMREADY: update avg tested subsets to be from the ICB timeouts
	among those, ``Average tested subset SSes'' counts how many partial verifications \quicksand~provided on average for each test (\sect{\ref{sec:eval-sssmc}}).
	``Total DR PPs'' counts how many unique data-racing instructions we identified among tests where we found no bugs;
	among those, ``Untested DR PPs'' counts how many could not be checked in the time limit (\sect{\ref{sec:future}}).
		}
	\label{tab:drbugs}
\end{table*}

@@

{\bf Partial verification.}
When a MC job times out, the user may prefer a brief summary of what parts of the test were verified, rather than writing off all the CPU time as a waste.
%Beyond using the state space estimator to guess the percent coverage,
We know of no approach to quantify the probability
that a bug would be exposed by an untested interleaving,
but \quicksand~at least reports which subsets of PPs resulted in state spaces that did complete in time.
%\quicksand~does not attempt any such quantitative guarantee,
%but can at least report
On 129 tests, the SSS-MC control experiment timed out after 10 hours with no bugs found.
Among these tests, \quicksand~found bugs in 36.
For the other 93, we show the number of state spaces \quicksand~was able to complete in the ``Average tested subset SSes'' column of Table~\ref{tab:drbugs}.
%and the number of unique PPs among them.
%In 6 cases, \quicksand~also failed to complete anything; beyond these,
%between 1 and 253 state spaces were completed for each test
%Between 0 and 233 unique PPs were tested among some completed subsets, with a mean of 27.2 and median of 7.5.
These completions guarantee that, if the test program could expose a bug,
% Justify implicit hypothesis: Sync PPs + DR PPs = all possible relevant PPs.
it would only be found by a new data-race PP not discovered yet, or by a superset combination of PPs not reached.
\revision{
%While obviously not as strong as full verification,
Prior work \cite{sealing} has argued the value of similar {\em compositional testing} when full verification is intractable,
deferring to the user's expertise to judge the value of each subset of PPs verified.
}

@@

% TODO
% Actually, while SSS-MC-Shared-Mem proved competitive with \quicksand~at finding data-race-related bugs,
% its heuristic to ignore threads' own stack accesses introduces unsoundness.
% <Move contents of footnote here.>
% Indeed, SSS-MC-Shared-Mem claimed a full verification on 999999 tests for which \quicksand~found bugs!
% % Hence, no SSS-MC strategy competes with quicksand for truly sound verification
% Hence, only SSS-MC-DPOR and SSS-MC-ICB compete with \quicksand for sound verification, and they are thoroughly outperformed.
% Between data-race bugs and verifications, \quicksand~is shown to provide the best of both worlds.
